### 2017-02-23 paper/0_abstract.tex
paper: cleanup line breaks and break into separate files
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\begin{notes}
\textcolor{red}{
\noindent Issues raised by reviewers, paper augmented as noted inline (red/blue text):
\begin{enumerate}
  \item clarify what was implemented: diagram of interfaces and prose
  specifying the difference between Malacology and prior work (reviewers B, C,
  D, E)
  \item describe what makes a programmable storage system (reviewers C, D)
  \item organize motivation: group interfaces into categories (e.g., core
  functionality, features, and performance optimizations) and describe other
  applications we can build with Malacology (reviewers B, C)
  \item clarify terminology: service vs. file system metadata, storage vs.
  file system, dirty vs. clean slate approaches, soft-state caching (reviewers
  B, C)
  \item address security, access control, and the safety; specify how it
  affects composability, performance, and multi-tenancy (reviewer D)
\end{enumerate}
}
\end{notes}

\begin{abstract} Storage systems are caught between the need to evolve data
processing applications efficiently and quickly, and the increasing velocity
with which storage device technology evolves. This puts tremendous pressure on
storage systems to support rapid change both in terms of their interfaces and
their performance. But adapting storage systems can be difficult because
unprincipled changes might jeopardize years of code-hardening and performance
optimization efforts that were necessary for users to entrust their data to the
storage system. We introduce the concept of a programmable storage system, a
system that exposes internal services and abstractions with the goal of
re-using and building new services on top of the storage stack. We present
Malacology, a prototype implementation of a programmable storage system to
explore how existing abstractions of common storage system services can be
leveraged to adapt to the needs of new data processing systems and the
increasing variety of storage devices. We show how this approach allows a large
degree of flexibility for storage systems to evolve without sacrificing the
robustness of their code-hardened subsystems. We illustrate the advantages and
challenges of programmability by composing existing primitives into two new
higher-level services: a \newcommentone{file system} metadata load balancer and
a high-performance distributed shared-log. The evaluation demonstrates that our
services inherit desirable qualities of the back-end storage system, including
the ability to balance load, efficiently propagate cluster metadata, recover
from failure, and to navigate trade-offs between latency and throughput using
leases.  \addressesissue{4}
\end{abstract}