### 2017-02-23 paper/7_future_work.tex

\section{Future Work}

\newcommentone{This work is one point along the path towards producing
general-purpose storage systems that can target special-purpose applications.
Continuing to expore existing storage system functionality as reusable building
blocks will enable the proliferation of storage-centric APIs, allowing
developers to custom-fit their applications to take advantage of the
capabilities of the underlying system. When adding new interfaces and composing
them together, careful consideration should be made to address security, access
control, and safety. Correctness and performance will surely be affected if
interfaces are combined poorly and these drawbacks will be quantified.}
\addressesissue{5}

\newcommentone{The next higher-level services we plan to build using the
interfaces in Table~\ref{table:examples} are: an elastic cloud database, a data
layout manager, and a data processing engine. Some cloud databases aggressively
cache data to hide the overheads and weaker consistency semantics of the
distributed file systems they sit on~\cite{pavlo:sigmod16} but using Malacology
we could build a database that works well because of the storage system not in
spite of it. We plan to use the data IO interface for push down predicates, the
file type interface for maintaining indices and stats, and the durability
interface internals for ingest and data movement. The data layout manager would
be similar to DRepl~\cite{ionkov:msst2013} and would use the file type,
durability, and data I/O interface to manage different views of data. Finally,
the data processing engine would be similar to MapReduce and would use the data
I/O, shared resource, and load balancing interface to implement the resource
manager, computation engine, and storage platform in the Hadoop ecosystem.}

\newcommentone{ Our experience with Zlog and Mantle illustrates that the labor
of wrapping existing services in interfaces is justified by the power and
flexibility they afford to programmers.  In exchange for this flexibility,
however, programmable storage systems force programmers to forfeit the
\emph{protection from change} afforded by narrow storage interfaces like the
POSIX API.  To implement applications built atop systems such as Malacology,
programmers must navigate a complex design space, simultaneously addressing
often orthogonal concerns including functional correctness, performance and
fault tolerance. Worse still, these compositions can be sensitive to changes in
the underlying environment (such as hardware upgrades, software version changes
and evolving workloads).  For example, a major version change in Ceph required
us to rewrite significant parts of Zlog to maintain acceptable performance --
each such evolution costs developer time and risks introducing bugs.}

\newcommentone{We are actively exploring the use of high-level
\emph{declarative} languages based on Datalog~\cite{alvaro:cidr11} to program
storage APIs.  In this way, a programmer can specify the functional behavior of
a storage interface once in a relational (or algebraic) language, allowing an
optimizer to model search the space of functionally equivalent physical
implementations, and re-optimizing when characteristics change.  Much like
query planning and optimization in database systems~\cite{hellerstein:cidr15},
this approach will separate the concerns of correctness and performance, and
protect applications (which evolve slowly) against changes in highly dynamic
storage system.} \addressesissue{3}