### 2017-02-23 paper/9_conclusion.tex

\section{Conclusion}
\label{conclusion-and-future-work}

Programmable storage is a viable method for eliminating duplication of complex
error-prone software used as workarounds for storage system deficiencies. We
propose that systems expose their services in a safe way allowing application
developers to customize system behavior to meet their needs while not
sacrificing correctness. To illustrate the benefits of this approach we
presented Malacology, a programmable storage system that facilitates the
construction of new services by re-purposing existing subsystem abstractions of
the storage stack.

\oldcommentone{Future work will focus on constructing more interfaces to
support a wide variety of storage system services that can be configured
on-the-fly in existing systems. This work is one point along that path to
producing general-purpose storage systems that can target  special-purpose
applications.  Ultimately we want to utilize declarative methods for expressing
new services.}