### 2017-02-23 paper/paper.tex

\documentclass[preprint]{sigplanconf-eurosys}
\usepackage{multirow}
\usepackage{arydshln}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{pifont}
\usepackage{color}
\usepackage{comment}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Uncomment for diff 0
%\includecomment{notes}
%\newcommand{\newcomment}[1]{{\textcolor{red}{\textbf{#1}}}}
%\newcommand{\oldcomment}[1]{{\textcolor{blue}{\textbf{\sout{#1}}}}}
%\newcommand{\addressesissue}[1]{{\textcolor{red}{{\it (Addresses issue(s): {#1})}}}}

% Uncomment for diff 1
%\includecomment{notes}
%\newcommand{\newcommentone}[1]{{\textcolor{red}{\textbf{#1}}}}
%\newcommand{\oldcommentone}[1]{{\textcolor{blue}{\textbf{\sout{#1}}}}}
%\newcommand{\newcomment}[1]{{\textcolor{black}{#1}}}
%\newcommand{\oldcomment}[1]{}
%\newcommand{\addressesissue}[1]{{\textcolor{red}{{\it (Addresses issue(s): {#1})}}}}

% Uncomment for camera-ready
\excludecomment{notes}
\newcommand{\newcomment}[1]{{\textcolor{black}{#1}}}
\newcommand{\oldcomment}[1]{}
\newcommand{\newcommentone}[1]{{\textcolor{black}{#1}}}
\newcommand{\oldcommentone}[1]{}
\newcommand{\addressesissue}[1]{}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\cmark}{\ding{51}}%
\usepackage[font=footnotesize,labelfont=bf]{caption}

\setlength{\belowcaptionskip}{-10pt}

\newcommand{\paa}[1]{{\textcolor{red}{[[#1 -- paa]]}}}
\usepackage[normalem]{ulem}
\usepackage{pstricks}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
\setkeys{Gin}{width=\maxwidth,keepaspectratio}

\usepackage[unicode=true]{hyperref}
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={Michael A. Sevilla, Noah Watkins, Ivo Jimenez, Carlos Maltzahn, Peter Alvaro, Shel Finkelstein, Jeff LeFevre},
            pdftitle={Malacology: A Programmable Storage System},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=black,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\title{Malacology: A Programmable Storage System}
%\authorinfo{Paper ID: 165}          {Total Pages: 12}
\authorinfo{Michael A. Sevilla, Noah Watkins, Ivo Jimenez, \\Peter Alvaro, Shel Finkelstein, Jeff LeFevre, Carlos Maltzahn}
           {\{msevilla, jayhawk, ivo\}@soe.ucsc.edu, \{palvaro, shel, jlefevre, carlosm\}@ucsc.edu}
\date{}
\begin{document}

  \maketitle

  \input{./0_abstract}
  \input{./1_intro}
  \input{./2_application_specific_storage_stacks}
  \input{./3_challenges}
  \input{./4_malacology_a_programmable_storage_system}
  \input{./5_services_built_on_malacology}
  \input{./6_evaluation}
  \input{./7_future_work}
  \input{./8_related_work}
  \input{./9_conclusion}

\bibliography{paper}
\bibliographystyle{plain}

\end{document}