### 2017-03-14 paper/paper.tex

%-----------------------------------------------------------------------------
%
%               Template for sigplanconf LaTeX Class
%
% Name:         sigplanconf-template.tex
%
% Purpose:      A template for sigplanconf.cls, which is a LaTeX 2e class
%               file for SIGPLAN conference proceedings.
%
% Guide:        Refer to "Author's Guide to the ACM SIGPLAN Class,"
%               sigplanconf-guide.pdf
%
% Author:       Paul C. Anagnostopoulos
%               Windfall Software
%               978 371-2316
%               paul@windfall.com
%
% Created:      15 February 2005
% Source:       http://www.sigplan.org/Resources/Author/
%-----------------------------------------------------------------------------

\documentclass[numbers]{sigplanconf}

% The following \documentclass options may be useful:

% preprint       Remove this option only once the paper is in final form.
%  9pt           Set paper in  9-point type (instead of default 10-point)
% 11pt           Set paper in 11-point type (instead of default 10-point).
% numbers        Produce numeric citations with natbib (instead of default author/year).
% authorversion  Prepare an author version, with appropriate copyright-space text.

\usepackage{amsmath}
\usepackage{multirow}
\usepackage{arydshln}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{pifont}
\usepackage{color}
\usepackage{comment}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Uncomment for diff 0
%\includecomment{notes}
%\newcommand{\newcomment}[1]{{\textcolor{red}{\textbf{#1}}}}
%\newcommand{\oldcomment}[1]{{\textcolor{blue}{\textbf{\sout{#1}}}}}
%\newcommand{\addressesissue}[1]{{\textcolor{red}{{\it (Addresses issue(s): {#1})}}}}

% Uncomment for diff 1
%\includecomment{notes}
%\newcommand{\newcommentone}[1]{{\textcolor{red}{\textbf{#1}}}}
%\newcommand{\oldcommentone}[1]{{\textcolor{blue}{\textbf{\sout{#1}}}}}
%\newcommand{\newcomment}[1]{{\textcolor{black}{#1}}}
%\newcommand{\oldcomment}[1]{}
%\newcommand{\addressesissue}[1]{{\textcolor{red}{{\it (Addresses issue(s): {#1})}}}}

% Uncomment for camera-ready
\excludecomment{notes}
\newcommand{\newcomment}[1]{{\textcolor{black}{#1}}}
\newcommand{\oldcomment}[1]{}
\newcommand{\newcommentone}[1]{{\textcolor{black}{#1}}}
\newcommand{\oldcommentone}[1]{}
\newcommand{\addressesissue}[1]{}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\cmark}{\ding{51}}%
\usepackage[font=footnotesize,labelfont=bf]{caption}

\setlength{\belowcaptionskip}{-10pt}

\newcommand{\paa}[1]{{\textcolor{red}{[[#1 -- paa]]}}}
\usepackage[normalem]{ulem}
\usepackage{pstricks}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
\setkeys{Gin}{width=\maxwidth,keepaspectratio}

\usepackage[unicode=true]{hyperref}
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={Michael A. Sevilla, Noah Watkins, Ivo Jimenez, Carlos Maltzahn, Peter Alvaro, Shel Finkelstein, Jeff LeFevre},
            pdftitle={Malacology: A Programmable Storage System},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=black,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\newcommand{\cL}{{\cal L}}

\begin{document}

\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

\CopyrightYear{2017} 
\setcopyright{acmcopyright}
\conferenceinfo{EuroSys '17,}{April 23-26, 2017, Belgrade, Serbia}
\isbn{978-1-4503-4938-3/17/04}\acmPrice{\$15.00}
\doi{http://dx.doi.org/10.1145/3064176.3064208}

%\conferenceinfo{CONF'yy}{Month d--d, 20yy, City, ST, Country}
%\copyrightyear{20yy}
%\copyrightdata{978-1-nnnn-nnnn-n/yy/mm}\reprintprice{\$15.00}
%\copyrightdoi{nnnnnnn.nnnnnnn}

% For compatibility with auto-generated ACM eRights management
% instructions, the following alternate commands are also supported.
%\CopyrightYear{2016}
%\conferenceinfo{CONF'yy,}{Month d--d, 20yy, City, ST, Country}
%\isbn{978-1-nnnn-nnnn-n/yy/mm}\acmPrice{\$15.00}
%\doi{http://dx.doi.org/10.1145/nnnnnnn.nnnnnnn}

% Uncomment the publication rights used.
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}  % default
%\setcopyright{rightsretained}

\titlebanner{This is a draft!}        % These are ignored unless
\preprintfooter{short description of paper}   % 'preprint' option specified.

\title{Malacology: A Programmable Storage System}

\authorinfo{Michael A. Sevilla, Noah Watkins, Ivo Jimenez, \\Peter Alvaro, Shel Finkelstein, Jeff LeFevre, Carlos Maltzahn}
           {University of California, Santa Cruz}
           {\{msevilla, jayhawk, ivo\}@soe.ucsc.edu, \{palvaro, shel\}@ucsc.edu, \{jlefevre, carlosm\}@soe.ucsc.edu}

\maketitle

\input{./0_abstract}

% 2012 ACM Computing Classification System (CSS) concepts
% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10002951.10003152.10003517.10003519</concept_id>
<concept_desc>Information systems~Distributed storage</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10011007.10010940.10010941.10010949.10003512</concept_id>
<concept_desc>Software and its engineering~File systems management</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10011007.10010940.10010992</concept_id>
<concept_desc>Software and its engineering~Software functional properties</concept_desc>
<concept_significance>300</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Information systems~Distributed storage}
\ccsdesc[500]{Software and its engineering~File systems management}
\ccsdesc[300]{Software and its engineering~Software functional properties}
% end generated code

% Legacy 1998 ACM Computing Classification System categories are also
% supported, but not recommended.
%\category{CR-number}{subcategory}{third-level}[fourth-level]
%\category{D.3.0}{Programming Languages}{General}
%\category{F.3.2}{Logics and Meanings of Programs}{Semantics of Programming Languages}[Program analysis]

\keywords
Distributed Storage, Programmability, Ceph

\input{./1_intro}
\input{./2_application_specific_storage_stacks}
\input{./3_challenges}
\input{./4_malacology_a_programmable_storage_system}
\input{./5_services_built_on_malacology}
\input{./6_evaluation}
\input{./7_future_work}
\input{./8_related_work}
\input{./9_conclusion}

% The 'abbrvnat' bibliography style is recommended.
\bibliographystyle{abbrvnat}

% The bibliography should be embedded for final submission.
\bibliography{paper}

\end{document}